A user-extensible and adaptable parser architecture

### 2008-05-29 papers/sgai-2008/paper.tex
Removed a lot of content to shorten the paper.

@@

XXX DO I NEED TO SAY THAT THINGS WILL BE EXPANDED UPON IN MY THESIS\@?

@@

    \item The mail administrator can define custom rejection messages,
        resulting in custom log lines~\cite{postfix-lookup-tables,
        smtpd_access_readme, smtpd_per_user_control}.

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.\footnote{A review of the existing parsers
considered for this project will be included in the author's thesis.} The
effort required to adapt and improve an existing parser was judged to be
greater than the effort to write one from scratch, as the techniques used
by the existing parsers severely limited their potential: some ignored the
majority of log lines, parsing specific log lines with great accuracy, but
in an inextensible fashion; others sloppily parsed the majority of log
lines, but were incapable of distinguishing between log lines of the same
type, e.g.\ rejections, and extracted only a portion of the data required
for this project.

@@

The function which finds the rule matching the log line and invokes the
requested action can be expressed in pseudo-code (with indentation denoting
flow of control) as:

\begin{verbatim}

for each line in the input files:
    for each rule defined by the user:
        if this rule matches the input line:
            perform the action specified by the rule
            skip the remaining rules
            process the next input line
    warn the user that the input line was not parsed

\end{verbatim}

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches, or log lines lacking a way to identify which data structure
they should be associated with.\footnote{Postfix allocates a queueid
(explained in the glossary,~\sectionref{Glossary}) to each mail; most log
lines contain the queueid of the associated mail, but some log lines
(particularly with older versions of Postfix) lack a queueid, preventing
their association with a particular data structure.}  The parser warns
about every unparsed log line, to alert the user that they need to alter or
extend their rules; the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The remaining actions have only one or two associated rules: in some cases
this is because there will only ever be one type of log line for that
action, e.g.\ all log lines showing that a remote client has connected are
matched by a single rule and handled by the \texttt{CONNECT} action; in
other cases the action is required to address a deficiency in the log
files, or a complication which arises during
parsing.\footnote{\label{complications footnote}The complications which
arose during development of the parser will be documented in detail in the
author's thesis.}

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.\footref{complications footnote}  Some were the result of
deficiencies in Postfix's logging, which were improved or resolved by later
versions of Postfix; others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in the parser: adding new actions was enough for several of
the complications; others required modifications to a single existing
action to work around the difficulties; the remainder were resolved by
adapting actions to pass information to subsequent actions (via the
framework), which then check for the extra information and change their
behaviour appropriately.

@@

    \item [regex] The \regex{} to match the log line against.  For
        efficiency every rule's \regex{} is compiled before attempting to
        parse the log file --- otherwise each \regex{} would be recompiled
        each time it was used, resulting in a large, data dependent
        slowdown.\footnote{This will be discussed in more detail in the
        author's thesis.}  XXX MAYBE SCRAP THE SECOND SENTENCE\@?

@@

XXX MAYBE REMOVE THE SECTION ON DETECTING OVERLAPPING RULES\@?

Detecting overlapping rules is difficult, but the following approaches may
be helpful:

    \item Sort by \regex{} and visually inspect the list, e.g.\ with \SQL{}
        similar to: \textbf{select regex from rules order by regex;}

    \item Compare the results of parsing using sorted, shuffled and
        reversed rules.\footnote{See \sectionref{rule efficiency} for more
        details of sorting the rules.}  Parse a number of log files using
        normal sorting, then dump a textual representation of the rules,
        connections and results tables.  Repeat with shuffled and reversed
        rules, starting with a fresh database.  If there are no overlapping
        rules the tables from each run will be identical; differences
        indicate overlapping rules.  Which rules overlap can be determined
        by examining the differences in the tables: each result contains a
        reference to the rule which created it, if the references differ
        between runs the two rules referenced in the differing records
        overlap.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the input files which match more than one
        rule.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would produce much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the results from first run discarded,
and the execution time for the remaining 9 runs averaged.  XXX DO I NEED TO
EXPLAIN WHY\@?  The first run is discarded for two reasons:

\begin{enumerate}

    \item The execution time will be higher because the log file must be
        read from disk, whereas for subsequent runs the log file will be
        cached in memory by the operating system.

    \item The execution time will also be higher because the rule ordering
        will be sub-optimal compared to subsequent runs.

\end{enumerate}

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB, and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time, and ratio of number of log lines vs execution time (higher
is better in both cases).  As the reader can see the ratios are quite
tightly banded across the graph, showing that the algorithm scales
linearly: the much larger log files at points 62--68 on the X axis in
fig.~\refwithpage{execution time vs file size vs number of lines graph}
actually cause the ratio to increase (i.e.\ improve), rather than decrease.
The strange behaviour where larger log files are parsed more efficiently is
explained in \sectionref{Why are there dips in the graphs?}.

\showgraph{build/plot-normal-filesize-numlines}{Execution time vs file
size vs number of lines}{execution time vs file size vs number of lines
graph}

\showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file
size and number of lines to execution time}{execution time vs file size vs
number lines factor}

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{execution time vs file size vs number of
lines graph}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor} (where a peak means that more lines are processed per
second, i.e.\ performance is better).  The explanation for this took some
time to arrive at, but it turns out to be reasonably simple.  The large log
files in question were caused by a mail forwarding loop, where the
distribution of log lines is quite different to normal, resulting in
different performance characteristics.\footnote{The cause of the mail
forwarding loop will be explained in the author's thesis.}

@@

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.\footnote{There will be substantially more
coverage discussion in the author's thesis.}  The former is initially more
important, as the parser must successfully parse every line if it is to be
complete, but subsequently the latter takes precedence because reproducing
the path a mail takes through Postfix is the aim of the parser.  Increasing
the percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

>>>>

\today{}

@@

@@

    \item The mail administrator can define custom rejection
        messages~\cite{postfix-lookup-tables, smtpd_access_readme,
        smtpd_per_user_control}, resulting in custom log lines.

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.  The effort required to adapt and improve an
existing parser was judged to be greater than the effort to write one from
scratch, as the techniques used by the existing parsers severely limited
their potential: some ignored the majority of log lines, parsing specific
log lines with great accuracy, but in an inextensible fashion; others
sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines of the same type, e.g.\ rejections, and
extracted only a portion of the data required for this project.

@@

The function which finds the rule matching the log line and invokes the
requested action can be expressed in pseudo-code (with indentation denoting
flow of control):

% DO NOT REFORMAT!

\begin{verbatim}
for each line in the input files: 
    for each rule defined by the user: 
        if this rule matches the input line:
            perform the action specified by the rule
            skip the remaining rules
            process the next input line
    warn the user that the input line was not parsed
\end{verbatim}

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches, or log lines lacking a way to identify which data structure
they should be associated with.\footnote{Postfix allocates a queueid
(explained in the glossary,~\sectionref{Glossary}) to each mail; most log
lines contain the queueid of the associated mail, but some log lines
(particularly with older versions of Postfix) lack a queueid, preventing
their association with a particular data structure.}  The parser warns
about every unparsed log line, to alert the user that they need to alter or
extend their rules; the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The remaining actions have only one or two associated rules: in some cases
this is because there will only ever be one type of log line for that
action, e.g.\ all log lines showing that a remote client has connected are
matched by a single rule and handled by the \texttt{CONNECT} action; in
other cases the action is required to address a deficiency in the log
files, or a complication which arises during parsing.

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.  Some were the result of deficiencies in Postfix's logging,
which were improved or resolved by later versions of Postfix; others were
due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the parser:
adding new actions was enough for several of the complications; others
required modifications to a single existing action to work around the
difficulties; the remainder were resolved by adapting actions to pass
information to subsequent actions (via the framework), which then check for
the extra information and change their behaviour appropriately.

@@

    \item [regex] The \regex{} to match the log line against.

@@

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would produce much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the results from first run discarded,
and the execution time for the remaining 9 runs averaged.

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB, and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time, and ratio of number of log lines vs execution time (higher
is better in both cases).  As the reader can see the ratios are quite
tightly banded across the graph, showing that the algorithm scales
linearly: the much larger log files at points 62--68 on the X axis in
fig.~\refwithpage{execution time vs file size vs number of lines graph}
actually cause the ratio to increase (i.e.\ improve), rather than decrease.
The strange behaviour where larger log files are parsed more efficiently is
explained in \sectionref{Why are there dips in the graphs?}.
\showgraph{build/plot-normal-filesize-numlines}{Execution time vs file size
vs number of lines}{execution time vs file size vs number of lines graph}
\showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file size
and number of lines to execution time}{execution time vs file size vs
number lines factor}

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{execution time vs file size vs number of
lines graph}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor} (where a peak means that more lines are processed per
second, i.e.\ performance is better).  The explanation for this took some
time to arrive at, but it turns out to be reasonably simple.  The large log
files in question were caused by a mail forwarding loop, where the
distribution of log lines is quite different to normal, resulting in
different performance characteristics.

@@

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.
